
# Reproducible Research | Course Project 1
===========================================


### 1. Loading and preprocessing the data
#### 1.1. Set up working directory and load the data

*Set the correct R working directory with the setwd() function.*

```{r}
activityData <- read.csv('activity.csv', stringsAsFactors=FALSE)
```

#### 1.2 Process/transform the data into a format suitable for the analysis

```{r}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```


### 2. What is mean total number of steps taken per day?
*For this part of the assignment, you can ignore the missing values in the dataset.*

#### 2.1. Calculate the total number of steps taken per day

```{r}
StepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```

#### 2.2. Make a histogram of the total number of steps taken each day

```{r}
hist(StepsByDay, 
     col="red", 
     breaks=seq(from=0, to=25000, by=2500),
     xlab="Total number of steps", 
     ylim=c(0, 20), 
     main="Total number of steps taken each day")
```

#### 2.3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
StepsByDayMean <- mean(StepsByDay)
StepsByDayMedian <- median(StepsByDay)
```
* Mean: `r StepsByDayMean`
* Median:  `r StepsByDayMedian`


### 3. What is the average daily activity pattern?

#### 3.1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
AverageSteps <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

```{r}
library(ggplot2)
ggplot(data=AverageSteps, aes(x=interval, y=meanSteps)) +
    geom_line(color='red')+
    xlab("5-minute interval")+
    ylab("Average number of steps taken") 
```

#### 3.2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
MaxSteps <- which(AverageSteps$meanSteps == max(AverageSteps$meanSteps))
TimeMaxSteps <- AverageSteps[MaxSteps, 1]
```

* Maximum number of steps at: `r TimeMaxSteps`


### 4. Imputing missing values

#### 4.1. Calculate and report the total number of missing values in the dataset.

```{r}
MissingValues <- length(which(is.na(activityData$steps)))
```

* Total number of missing values: `r MissingValues`

#### 4.2. Devise a strategy for filling in all of the missing values in the dataset. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
library(Hmisc)
ActivityDataImp <- activityData
ActivityDataImp$steps <- impute(activityData$steps, fun=mean)
```

#### 4.4. Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps? 

```{r}
StepsByDayImp <- tapply(ActivityDataImp$steps, ActivityDataImp$date, sum)
hist(StepsByDayImp, 
     breaks=seq(from=0, to=25000, by=2500),
     col="red", 
     xlab="Total number of steps", 
     ylim=c(0, 30), 
     main="Total number of steps taken each day (missing data imputed)")
```

#### Mean and median total number of steps taken per day

```{r}
StepsByDayMeanImp <- mean(StepsByDayImp)
StepsByDayMedianImp <- median(StepsByDayImp)
```

* Mean (Imputed): `r StepsByDayMeanImp`
* Median (Imputed):  `r StepsByDayMedianImp`

*The values differ  from the estimates from the first part of the assignment. By adding the missing values, there is more data and therefore the mean and media values are bigger.*


### 5. Are there differences in activity patterns between weekdays and weekends?
* For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.*

#### 5.1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
ActivityDataImp$dateType <-  ifelse(as.POSIXlt(ActivityDataImp$date)$wday %in% c(0,6), 'Weekend', 'Weekday')
```

#### 5.2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
AvgActivityDataImp <- aggregate(steps ~ interval + dateType, data=ActivityDataImp, mean)
ggplot(AvgActivityDataImp, aes(interval, steps)) + 
    geom_line(color='red') + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("Average number of steps")
```
